
```{r}
library(car)
library(GGally)

```


READING IN THE DATA FROM THE CSV AND ONLY SELECTING THE PREDICTOR VARIABLES WE PLAN TO LOOK AT FOR OUR GROUP PROJECT
```{r}
df = read.csv('https://raw.githubusercontent.com/ethan2411/Data-603-604/main/603%20Data/500_Cities__Census_Tract-level_Data__GIS_Friendly_Format___2019_release_20231106.csv')


df.calcs <- subset(df, select =- c(StateAbbr, PlaceName, PlaceFIPS, TractFIPS,Place_TractID,Population2010, ACCESS2_Crude95CI, ARTHRITIS_Crude95CI, BINGE_Crude95CI, BPHIGH_Crude95CI, BPMED_Crude95CI, CANCER_Crude95CI, CASTHMA_Crude95CI, CHD_Crude95CI, CHECKUP_Crude95CI, CHOLSCREEN_Crude95CI, COLON_SCREEN_Crude95CI,COPD_Crude95CI, COREM_Crude95CI, COREW_Crude95CI, CSMOKING_Crude95CI, DENTAL_Crude95CI,DIABETES_Crude95CI,  HIGHCHOL_Crude95CI, KIDNEY_Crude95CI, LPA_Crude95CI, MAMMOUSE_Crude95CI, MHLTH_Crude95CI, OBESITY_Crude95CI, PAPTEST_Crude95CI, PHLTH_Crude95CI, SLEEP_Crude95CI, STROKE_Crude95CI, TEETHLOST_Crude95CI, Geolocation))

df.unhealthy.precitors <- df.calcs <- subset(df.calcs, select =- c(ARTHRITIS_CrudePrev, BPHIGH_CrudePrev, CASTHMA_CrudePrev, CHD_CrudePrev, COPD_CrudePrev, DIABETES_CrudePrev, HIGHCHOL_CrudePrev, KIDNEY_CrudePrev, MAMMOUSE_CrudePrev, MHLTH_CrudePrev, PHLTH_CrudePrev, STROKE_CrudePrev))

cancer.pred.only <- subset(df.unhealthy.precitors,  select =- c(BINGE_CrudePrev, CSMOKING_CrudePrev, LPA_CrudePrev, OBESITY_CrudePrev, SLEEP_CrudePrev))
cancer.df = na.omit(data.frame(cancer.pred.only))
head(cancer.df)

```



MAKE THE FIRST LINEAR MODEL FOR THE PROJECT

```{r}
cancer.lm = lm(CANCER_CrudePrev ~ ., data = cancer.df)
summary(cancer.lm)
```


RUN VIF to see which variables can be dropped to avoid multicollinearity then build the model. WILL DROP ANY ABOVE 5


```{r}

vif(cancer.lm)

```

DROP HIGHEST VIF (DENTAL_CRUDEPrev) THEN TEST VIF AGAIN

```{r}
cancer.lm2 = lm(CANCER_CrudePrev ~ ACCESS2_CrudePrev + BPMED_CrudePrev + CHECKUP_CrudePrev + CHOLSCREEN_CrudePrev + COLON_SCREEN_CrudePrev + COREM_CrudePrev + COREW_CrudePrev + PAPTEST_CrudePrev + TEETHLOST_CrudePrev, data = cancer.df)

vif(cancer.lm2)

```

DROP COLON_SCREEN_CRUDEPREV


```{r}
cancer.lm3 = lm(CANCER_CrudePrev ~ ACCESS2_CrudePrev + BPMED_CrudePrev + CHECKUP_CrudePrev + CHOLSCREEN_CrudePrev + COREM_CrudePrev + COREW_CrudePrev + PAPTEST_CrudePrev + TEETHLOST_CrudePrev, data = cancer.df)

vif(cancer.lm3)
```

DROP TEETHLOST_CrudePREV

```{r}
cancer.lm4 = lm(CANCER_CrudePrev ~ ACCESS2_CrudePrev + BPMED_CrudePrev + CHECKUP_CrudePrev + CHOLSCREEN_CrudePrev + COREM_CrudePrev + COREW_CrudePrev + PAPTEST_CrudePrev, data = cancer.df)

vif(cancer.lm4)
```

NO MORE TO DROP. USE THIS MODEL

NOW USE t-test TO SEE IF YOU CAN DROP ANY VARIABLES FROM THE 7 VARIABLE MODEL

```{r}
summary(cancer.lm4)
```

WE SEE FROM THE SUMMARY WE DONT NEED TO DROP ANY VARIABLES BECAUSE ALL < 0.05


NOW ADD INTERACTIONS 


```{r}
cancerfullint = lm(CANCER_CrudePrev ~ (ACCESS2_CrudePrev + BPMED_CrudePrev + CHECKUP_CrudePrev + CHOLSCREEN_CrudePrev + COREM_CrudePrev + COREW_CrudePrev + PAPTEST_CrudePrev)^2, data = cancer.df)

summary(cancerfullint)

anova(cancerfullint, cancer.lm4)

```

WE ALSO SEE FROM THE PARTIAL F TEST THAT WE CAN ACCEPT THE INTERACTION MODE BECAUSE P-VALUE < 0.05


FROM T-TEST VALUES WE DROP INTERACTION TERMS ACCESS2_CrudePrev:COREM_CrudePrev, BPMED_CrudePrev:COREM_CrudePrev, CHECKUP_CrudePrev:COREW_CrudePrev


NOW WE CHECK WHAT INTERACTIONS WE CAN DROP USING T-TEST. WE WILL DROP ALL OF THEM AT ONCE SO WE WILL USE PARTIAL F TEST TO SEE IF WE CAN ACCEPT REDUCED INT MODEL
```{r}

cancerintreduced = lm(CANCER_CrudePrev ~ ACCESS2_CrudePrev + BPMED_CrudePrev + CHECKUP_CrudePrev + CHOLSCREEN_CrudePrev + COREM_CrudePrev + COREW_CrudePrev + PAPTEST_CrudePrev + ACCESS2_CrudePrev:BPMED_CrudePrev + ACCESS2_CrudePrev:CHECKUP_CrudePrev + ACCESS2_CrudePrev:CHOLSCREEN_CrudePrev + ACCESS2_CrudePrev:COREW_CrudePrev + ACCESS2_CrudePrev:PAPTEST_CrudePrev + BPMED_CrudePrev:CHECKUP_CrudePrev + BPMED_CrudePrev:CHOLSCREEN_CrudePrev + BPMED_CrudePrev:COREW_CrudePrev + BPMED_CrudePrev:PAPTEST_CrudePrev + CHECKUP_CrudePrev:CHOLSCREEN_CrudePrev + CHECKUP_CrudePrev:COREM_CrudePrev + CHECKUP_CrudePrev:PAPTEST_CrudePrev + CHOLSCREEN_CrudePrev:COREM_CrudePrev + CHOLSCREEN_CrudePrev:COREW_CrudePrev + CHOLSCREEN_CrudePrev:PAPTEST_CrudePrev + COREM_CrudePrev:COREW_CrudePrev + COREM_CrudePrev:PAPTEST_CrudePrev + COREW_CrudePrev:PAPTEST_CrudePrev, data = cancer.df)

summary(cancerintreduced)

anova(cancerfullint,cancerintreduced)
```

ANOVA LETS US ACCEPT REDUCED INT MODEL



NOW ONTO MULTIPLE REGRESSIONS ASSUMPTIONS 


We will first start with linearity assumption and if we do not meet this assumption we will add higher powers

LINEARITY:

We check to see if there is linearity

```{r}
plot(cancerintreduced, which = 1)

```


Now we check the plots of the data to see where relationships are present and what powers we need to add

```{r}
ggpairs(cancerintreduced, lower = list(continuous = "smooth_loess", combo = "facethist", discrete = "facetbar", na = "na"))

```


Independence:

WE ASSUME INDEPENDENCE IS GOOD SO NO NEED TO TEST

Equal Variance:

```{r}
#residual vs fitted plot and scale location plot
plot(cancerintreduced, which = 1)
plot(cancerintreduced, which = 3)

```

test for normality
```{r}

bptest(cancerintreduced)

```



Normality:

histogram:
```{r}
ggplot(data=cancer.df, aes(residuals(cancerintreduced))) + 
  geom_histogram(breaks = seq(-20,20,by=0.1), col="green3", fill="green4") + 
  labs(title="Histogram for residuals") +
  labs(x="residuals", y="Count")
```


QQ plot
```{r}
plot(cancerintreduced, which = 2)
```


Shapiro Wilk test
```{r}
shapiro.test(cancerintreduced)
```


Multicollinearity:

double check multicolin of the variables w/out interactions
```{r}
vif(cancer.lm4)
```


Outliers:

Residual vs leverage plot
```{r}
plot(cancerintreduced, which = 5)
```


Summary of assumptions and what to do next:

Linearity assumption is ___ met


Normality assumption is ___ met


Equal variance assumption is ____ met

There are ____ outliers present


SO next, we will do _____









































































































































































