




```{r}
test.lm <- lm(Life.expectancy ~ factor(Country) + factor(Status) + Adult.Mortality + infant.deaths + Alcohol + percentage.expenditure + Hepatitis.B  + Measles + BMI + under.five.deaths + Polio + Total.expenditure + Diphtheria + HIV.AIDS + thinness.5.9.years+ Income.composition.of.resources + Schooling + GDP + Population + thinness..1.19.years, data = LE)
summary(test.lm)
```

```{r}
LE2 = subset(LE, select =- c(Year, Country))
LE2 = na.omit(LE2)
head(LE2)
```

```{r}
LE.lm <- lm(Life.expectancy ~ factor(Status) + Adult.Mortality + infant.deaths + Alcohol + percentage.expenditure + Hepatitis.B  + Measles + BMI + under.five.deaths + Polio + Total.expenditure + Diphtheria + HIV.AIDS + thinness.5.9.years+ Income.composition.of.resources + Schooling + GDP + Population + thinness..1.19.years, data = LE2)
summary(LE.lm)
```

REMOVING GDP, Population, and thinness 1-19 (to see if that's chill)
```{r}
LE.reduced <- lm(Life.expectancy ~ factor(Status) + Adult.Mortality + infant.deaths + Alcohol + percentage.expenditure + Hepatitis.B  + Measles + BMI + under.five.deaths + Polio + Total.expenditure + Diphtheria + HIV.AIDS + thinness.5.9.years+ Income.composition.of.resources + Schooling , data = LE2)
summary(LE.reduced)
anova(LE.lm, LE.reduced)
```
REMOVE MEASLES Hepatitis b, and Polio

```{r}
LE.reduced2 <- lm(Life.expectancy ~ factor(Status) + Adult.Mortality + infant.deaths + Alcohol + percentage.expenditure + BMI + under.five.deaths + Total.expenditure + Diphtheria + HIV.AIDS + thinness.5.9.years+ Income.composition.of.resources + Schooling , data = LE2)
summary(LE.reduced2)
anova(LE.reduced2, LE.reduced)
```
remove total expenditure 

```{r}
LE.reduced3 <- lm(Life.expectancy ~ factor(Status) + Adult.Mortality + infant.deaths + Alcohol + percentage.expenditure + BMI + under.five.deaths + Diphtheria + HIV.AIDS + thinness.5.9.years+ Income.composition.of.resources + Schooling , data = LE2)
summary(LE.reduced3)
anova(LE.reduced3, LE.reduced)
```
```{r}
plot(LE.reduced3)
```
```{r}
bptest(LE.reduced3)
shapiro.test(LE2$Life.expectancy)
```
```{r}
vif(LE.reduced3)
```

DROPPING THE FUCK OUT OF under.five.deaths

```{r}
LE.reduced4 <- lm(Life.expectancy ~ factor(Status) + Adult.Mortality + Alcohol + percentage.expenditure + BMI + infant.deaths + Diphtheria + HIV.AIDS + thinness.5.9.years+ Income.composition.of.resources + Schooling , data = LE2)
summary(LE.reduced4)
vif(LE.reduced4)
```

```{r}
LE.reduced5 <- lm(Life.expectancy ~ factor(Status) + Adult.Mortality + Alcohol + percentage.expenditure + BMI + infant.deaths + Diphtheria + HIV.AIDS + Income.composition.of.resources + Schooling , data = LE2)
summary(LE.reduced5)
```

```{r}
plot(LE.reduced5)
```

#Checking for higher order variables

```{r}
#making dataframe with only variables we're using for the regression. 
LE3 <- subset(LE2, select = c(Adult.Mortality, Alcohol, percentage.expenditure, BMI, infant.deaths, Diphtheria, HIV.AIDS, Income.composition.of.resources, Schooling, Life.expectancy))
```


```{r}
pairs(LE3)
```


AIDS looks to be higher order 

```{r}
LE.power <- lm(Life.expectancy ~ factor(Status)  + Adult.Mortality + Alcohol + percentage.expenditure + BMI + infant.deaths + Diphtheria + poly(HIV.AIDS,5) + Income.composition.of.resources + Schooling , data = LE2)
summary(LE.power)
bptest(LE.power)
plot(LE.power)
```

REMOVE ALCOHOL

```{r}
LE.power2 <- lm(Life.expectancy ~ factor(Status)  + Adult.Mortality + percentage.expenditure + BMI + infant.deaths + Diphtheria + poly(HIV.AIDS,5) + Income.composition.of.resources + Schooling , data = LE2)
summary(LE.power2)
bptest(LE.power2)
plot(LE.power2)
```
















#Making an interactions model

```{r}
LE.interactions <- lm(Life.expectancy ~ (factor(Status) + Adult.Mortality + Alcohol + percentage.expenditure + BMI + infant.deaths + Diphtheria + HIV.AIDS + Income.composition.of.resources + Schooling)^2, data = LE2)
summary(LE.interactions)
```


#Using stepwise to reduce the model
```{r}
model_step = step(LE.interactions, direction = "both")
summary(model_step)
```
```{r}
summary(model_step)
plot(model_step)
bptest(model_step)
```

#LOG transformation

```{r}
log.lm <- lm(log(Life.expectancy) ~ factor(Status) + Adult.Mortality + 
    Alcohol + percentage.expenditure + BMI + infant.deaths + 
    Diphtheria + HIV.AIDS + Income.composition.of.resources + 
    Schooling + factor(Status):Adult.Mortality + factor(Status):Alcohol + 
    factor(Status):percentage.expenditure + factor(Status):BMI + 
    factor(Status):Income.composition.of.resources + factor(Status):Schooling + 
    Adult.Mortality:Alcohol + Adult.Mortality:BMI + Adult.Mortality:infant.deaths + 
    Adult.Mortality:Diphtheria + Adult.Mortality:HIV.AIDS + Adult.Mortality:Income.composition.of.resources + 
    Adult.Mortality:Schooling + Alcohol:percentage.expenditure + 
    Alcohol:HIV.AIDS + Alcohol:Income.composition.of.resources + 
    Alcohol:Schooling + percentage.expenditure:infant.deaths + 
    percentage.expenditure:Income.composition.of.resources + 
    percentage.expenditure:Schooling + BMI:infant.deaths + BMI:Diphtheria + 
    BMI:Income.composition.of.resources + BMI:Schooling + infant.deaths:Diphtheria + 
    infant.deaths:Income.composition.of.resources + infant.deaths:Schooling + 
    Diphtheria:Income.composition.of.resources + Diphtheria:Schooling + 
    HIV.AIDS:Schooling, data = LE2)

summary(log.lm)
plot(log.lm)
bptest(log.lm)
```

```{r}
library(MASS)
bc = boxcox(model_step,lambda=seq(-1,1))

bestlambda=bc$x[which(bc$y==max(bc$y))]
bestlambda
```

```{r}
sqrt.lm <- lm(sqrt(Life.expectancy) ~ factor(Status) + Adult.Mortality + 
    Alcohol + percentage.expenditure + BMI + infant.deaths + 
    Diphtheria + HIV.AIDS + Income.composition.of.resources + 
    Schooling + factor(Status):Adult.Mortality + factor(Status):Alcohol + 
    factor(Status):percentage.expenditure + factor(Status):BMI + 
    factor(Status):Income.composition.of.resources + factor(Status):Schooling + 
    Adult.Mortality:Alcohol + Adult.Mortality:BMI + Adult.Mortality:infant.deaths + 
    Adult.Mortality:Diphtheria + Adult.Mortality:HIV.AIDS + Adult.Mortality:Income.composition.of.resources + 
    Adult.Mortality:Schooling + Alcohol:percentage.expenditure + 
    Alcohol:HIV.AIDS + Alcohol:Income.composition.of.resources + 
    Alcohol:Schooling + percentage.expenditure:infant.deaths + 
    percentage.expenditure:Income.composition.of.resources + 
    percentage.expenditure:Schooling + BMI:infant.deaths + BMI:Diphtheria + 
    BMI:Income.composition.of.resources + BMI:Schooling + infant.deaths:Diphtheria + 
    infant.deaths:Income.composition.of.resources + infant.deaths:Schooling + 
    Diphtheria:Income.composition.of.resources + Diphtheria:Schooling + 
    HIV.AIDS:Schooling, data = LE2)

summary(sqrt.lm)
plot(sqrt.lm)
bptest(sqrt.lm)
```








WORKING WITH THE HEALTH FACTORS ONLY


```{r}

test.lm = lm(Life.expectancy ~  factor(Status) + Alcohol + Hepatitis.B  + Measles + BMI + Polio + Diphtheria + HIV.AIDS, data = LE)
summary(test.lm)

```

```{r}

vif(test.lm)


```

```{r}

test.lm2 = lm(Life.expectancy ~  factor(Status) + Alcohol + Hepatitis.B + BMI + Polio + Diphtheria + HIV.AIDS, data = LE)
summary(test.lm2)


```

```{r}

test.lmint = lm(Life.expectancy ~ (Alcohol + Hepatitis.B + BMI + Polio + Diphtheria + HIV.AIDS)^2, data = LE)
summary(test.lmint)

stepmod = ols_step_both_p(test.lmint, pent = 0.1, prem = 0.3, details = TRUE)

test.lmintred = step(test.lmint, method = "BOTH")
```

```{r}



summary(stepmod$model)
summary(test.lmintred)



stepmod_higher = lm(Life.expectancy ~ HIV.AIDS + I(HIV.AIDS^2) + I(HIV.AIDS^3) + I(HIV.AIDS^4) + Hepatitis.B + Diphtheria + Polio + Alcohol + BMI + BMI:Polio + Polio:Diphtheria + BMI:Alcohol + HIV.AIDS:BMI + Polio:Alcohol + Hepatitis.B:Alcohol, data = LE)

bptest(stepmod_higher)
plot(stepmod_higher)


library(MASS)
bc = boxcox(stepmod_higher, lambda=seq(-1,1))

bestlambda = bc$x[which(bc$y==max(bc$y))]
bestlambda

#ggpairs(stepmod_model)
#vif(test.lmintred)


```

```{r}

bcmodel=lm((((Life.expectancy^0.8383838)-1)/0.8383838) ~ HIV.AIDS + I(HIV.AIDS^2) + I(HIV.AIDS^3) + I(HIV.AIDS^4) + Hepatitis.B + Diphtheria + Polio + Alcohol + BMI + BMI:Polio + Polio:Diphtheria + BMI:Alcohol + HIV.AIDS:BMI + Polio:Alcohol + Hepatitis.B:Alcohol, data = LE)
summary(bcmodel)


bptest(bcmodel)
plot(bcmodel, which = 4)

```









```{r}



summary(test.lmintred)


```

```{r}
test.lmhigher = lm(Life.expectancy ~  factor(Status) + Alcohol + Hepatitis.B + BMI + Polio + Diphtheria + HIV.AIDS + I(HIV.AIDS^2) + I(HIV.AIDS^3) + I(HIV.AIDS^4) + factor(Status):Alcohol + factor(Status):Hepatitis.B + factor(Status):BMI + factor(Status):Diphtheria + Alcohol:Hepatitis.B + Alcohol:Diphtheria + Alcohol:Hepatitis.B + Alcohol:Diphtheria +  Alcohol:HIV.AIDS + Hepatitis.B:BMI + Hepatitis.B:Diphtheria + BMI:Polio + BMI:Diphtheria + Polio:Diphtheria + Diphtheria:HIV.AIDS, data = LE) 

bptest(test.lmhigher)

anova(test.lmhigher, test.lmintred)
```




```{r}
ggpairs(test.lmintred)

```


```{r}
plot(test.lmhigher)

summary(test.lmhigher)

```

```{r}
library(MASS)
bc = boxcox(test.lmhigher, lambda=seq(-1,1))

bestlambda = bc$x[which(bc$y==max(bc$y))]
bestlambda


```


```{r}
lev=hatvalues(test.lmhigher)
p = length(coef(test.lmhigher))
n = nrow(LE)
outlier2p = lev[lev>(2*p/n)]
outlier3p = lev[lev>(3*p/n)]
print("h_I>3p/n, outliers are")
print(outlier2p)

rows_remove = which(lev>(2*p/n))
print(outlier2plist)

LE_no_outliers = LE[-rows_remove, ]
head(LE_no_outliers)

nrow(LE)
nrow(LE_no_outliers)
```

```{r}
test.lm2 = lm(Life.expectancy ~  factor(Status) + Alcohol + Hepatitis.B  + Measles + BMI + Polio + Diphtheria + HIV.AIDS, data = LE_no_outliers)
summary(test.lm2)



```

```{r}

test.lm2red = lm(Life.expectancy ~  factor(Status) + Alcohol + Hepatitis.B  + Measles + BMI + Diphtheria + HIV.AIDS, data = LE_no_outliers)
summary(test.lm2red)

```

```{r}
test.lm2int = lm(Life.expectancy ~  (factor(Status) + Alcohol + Hepatitis.B  + Measles + BMI + Diphtheria + HIV.AIDS)^2, data = LE_no_outliers)
summary(test.lm2int)

step_model2 = step(test.lm2int)
summary(step_model2)

```

```{r}

test.lm2higher = lm(Life.expectancy ~  factor(Status) + Alcohol + Hepatitis.B  + Measles + I(Measles^2) + BMI + Diphtheria + HIV.AIDS + I(HIV.AIDS^2) + I(HIV.AIDS^3) + factor(Status):BMI + factor(Status):Diphtheria + Alcohol:Hepatitis.B + Alcohol:Measles + Alcohol:Diphtheria + Hepatitis.B:Diphtheria + Measles:Diphtheria + Measles:HIV.AIDS + BMI:Diphtheria + Diphtheria:HIV.AIDS,  data = LE_no_outliers)

summary(test.lm2higher)

bptest(test.lm2higher)


shapiro.test(residuals(test.lm2higher))


plot(test.lm2higher)


bc = boxcox(test.lm2higher, lambda=seq(-1,1))

bestlambda = bc$x[which(bc$y==max(bc$y))]
bestlambda



```

```{r}
ggpairs(test.lm2higher)


```






DATA 603 Final Project R CODE

This file contains all of our R code for the project. It contains our model building process in the order of these steps:

1. Building our first/initial model and doing a full F test.
2. Checking for multicollinearity and removing variables which have a VIF > 5
3. Reducing the model using the individual t-test. Use partial F-test to confirm the reduced model is significant
4. Adding interactions to the model and using partial F-test to confirm the interaction model is significant
5. Reducing the interaction model via stepwise selection
5. Checking plots between response variable and predictor variables to determine if higher powers need to be present
6. Adding higher powers and using partial F-test to confirm the hgher power model is significiant

7. Checking the assumptions of the final model
8. Doing various techniques if the model does not meet assumptions (boxcox transformation and removal of outliers)
9. rebuilding the models from the techniques done above and check if the new model meets assumptions better vs the other model. 

10. Choosing the best final model within the scope of this course
11. Interpreting beta coefficients



First we loaded in our dataset, removed na values and filtered for only rows that were considered "developing" countries based off of the column Status (states if a row is a devoloped or non-developed country). We filtered for these rows since our project only focuses on developing countries. We also loaded in any libraries we planned on using.

```{r}
library(car)
library(lmtest)
library(GGally)
library(olsrr)
library(MASS)
library(dplyr)
```

```{r}
#read in original csv file and remove na values
LE = read.csv("https://raw.githubusercontent.com/ethan2411/Data-603-604/main/Life%20Expectancy%20Data.csv")
LE = na.omit(LE)
head(LE)
```


```{r}
#select for "developing" rows based on Status column
developing_LE <- LE %>%
  filter(Status == "Developing")
nrow(LE)
nrow(developing_LE)
head(developing_LE)
```

1. Building our first/initial model and doing a full F test.
```{r}

#make first model
develop_firstlm = lm(Life.expectancy ~  Alcohol + Hepatitis.B  + Measles + BMI + Polio + Diphtheria + HIV.AIDS, data = developing_LE)
summary(develop_firstlm)

#check vif for multicolinn to see if we need to drop any variables
vif(develop_firstlm)
#we dont need to drop any so we now reduce model

#reduce model via t-test
develop_lmred = lm(Life.expectancy ~  Alcohol + Hepatitis.B + BMI + Polio + Diphtheria + HIV.AIDS, data = developing_LE)
summary(develop_lmred)

#check if reduced model is accepted (it is) via partial f test
anova(develop_firstlm,develop_lmred)
```

ANOVA ACCEPTS REDUCED MODEL

OLS STEP MODEL
```{r}
#make full interaction model
develop_lmintfull = lm(Life.expectancy ~  (Alcohol + Hepatitis.B + BMI + Polio + Diphtheria + HIV.AIDS)^2, data = developing_LE)
summary(develop_lmintfull)


#reduce interaction model via ols stepwise because there are too many interactions to do it manually
develop_stepmod_intred = ols_step_both_p(develop_lmintfull, pent= 0.1, prem = 0.3, details = TRUE)
summary(develop_stepmod_intred$model)

```

GGPAIRS FOR OLS STEP MODEL
```{r}

#code in the reduced interaction model from stepwise
developing_intred2 = lm(Life.expectancy ~ HIV.AIDS + Hepatitis.B + BMI + Alcohol + Polio + Diphtheria + Polio:Diphtheria +BMI:Polio + Hepatitis.B:Alcohol + HIV.AIDS:Alcohol + Diphtheria:Alcohol +  Hepatitis.B:Polio + HIV.AIDS:Diphtheria, data = developing_LE)

#look for graphs to see if higher order models are required
ggpairs(developing_intred2)


#ANOVA TO TEST IF REDUCED INTERACTION MODEL IS ACCEPTED. I DONT THINK WE NEED THIS AFTER DOING STEPWISE ANYWAYS 
anova(develop_firstlm, developing_intred2)

```

HIGHER ORDER OF OUR OLS() STEPMODEL
```{r}
developing_highermod = lm(Life.expectancy ~ HIV.AIDS + I(HIV.AIDS^2) + I(HIV.AIDS^3) + I(HIV.AIDS^4) + Hepatitis.B + BMI + Alcohol + Polio + Diphtheria + Polio:Diphtheria +BMI:Polio + Hepatitis.B:Alcohol + HIV.AIDS:Alcohol + Diphtheria:Alcohol +  Hepatitis.B:Polio + HIV.AIDS:Diphtheria, data = developing_LE)

#ANOVA TO TEST IF HIGHER ORDER MODEL IS ACCEPTED. IT IS ACCEPTED
anova(developing_highermod, developing_intred2)


#CHECK ASSUMPTIONS OF HIGHER ORDER MODEL
summary(developing_highermod)
bptest(developing_highermod)
plot(developing_highermod)


```


FROM ASSUMPTION CHECKS WE DO BOXCOX

```{r}
bc = boxcox(developing_highermod, lambda=seq(-1,1))

bestlambda = bc$x[which(bc$y==max(bc$y))]
bestlambda


```

boxcox is 1 so do nothing



OUTLIER CHECK WITH OLS_STEP MODEL SINCE WE ARE USING THIS STEPWISE REGRESSION


```{r}

lev=hatvalues(developing_highermod)
p = length(coef(developing_highermod))
n = nrow(developing_LE)
outlier2p = lev[lev>(2*p/n)]
outlier3p = lev[lev>(3*p/n)]
print("h_I>3p/n, outliers are")
print(outlier2p)

rows_remove3p = which(lev>(3*p/n))
print(rows_remove3p)

developing_LE_no_outliersSTEP = developing_LE[-rows_remove3p, ]
head(developing_LE_no_outliersSTEP)

nrow(developing_LE)
nrow(developing_LE_no_outliersSTEP)
```


TESTING FIT MODEL WITHOUT 3P OUTLIERS 
make new model then reduce
```{r}

first3plm= lm(Life.expectancy ~  Alcohol + Hepatitis.B  + Measles + BMI + Polio + Diphtheria + HIV.AIDS, data = developing_LE_no_outliersSTEP)
summary(first3plm)

#reduced model
red3plm = lm(Life.expectancy ~  Alcohol + Hepatitis.B + BMI + Polio + Diphtheria + HIV.AIDS, data = developing_LE_no_outliersSTEP)
summary(red3plm)

#int model
int3plm = lm(Life.expectancy ~  (Alcohol + Hepatitis.B + BMI + Polio + Diphtheria + HIV.AIDS)^2, data = developing_LE_no_outliersSTEP)
summary(int3plm)

intred3plm = ols_step_both_p(int3plm, pent = 0.1, prem = 0.3, details = FALSE)
summary(intred3plm$model)

#NEED TO MAKE THE REDUCED MODEL HERE AFTER RUNNING STEPWISE


```

```{r}

#MAKING REDUCED INT MODEL AFTER STEPWISE
step_red_3p = lm(Life.expectancy ~  HIV.AIDS + Hepatitis.B + Polio + Diphtheria + Alcohol + BMI + BMI:Diphtheria + Diphtheria:Polio + Hepatitis.B:Alcohol + Diphtheria:Alcohol + BMI:Hepatitis.B + BMI:Polio + HIV.AIDS:Polio + Diphtheria:Hepatitis.B, data = developing_LE_no_outliersSTEP)



#CHECK GGPAIRS TO SEE IF WE ADD HIGHER ORDER TO MODEL
#ggpairs(step_red_3p)

#LOOKS LIKE WE DO IT FOR HIV
#make higher order model
higher3plm = lm(Life.expectancy ~ HIV.AIDS + I(HIV.AIDS^2) + I(HIV.AIDS^3) + Hepatitis.B + Polio + Diphtheria + Alcohol + BMI + BMI:Diphtheria + Diphtheria:Polio + Hepatitis.B:Alcohol + Diphtheria:Alcohol + BMI:Hepatitis.B + BMI:Polio + HIV.AIDS:Polio + Diphtheria:Hepatitis.B, data = developing_LE_no_outliersSTEP)
#stopped at 3. 4 gives null values so i am avoiding that


#anova accepts higher order model
anova(higher3plm,step_red_3p)

summary(higher3plm)
bptest(higher3plm)
plot(higher3plm)

#shapiro.test(residuals(higher3plm))



```





NOW WE ATTEMPT WITH 2p OUTLIERS TO SEE IF ITS BETTER

```{r}
lev=hatvalues(developing_highermod)
p = length(coef(developing_highermod))
n = nrow(developing_LE)
outlier2p = lev[lev>(2*p/n)]
outlier3p = lev[lev>(3*p/n)]
print("h_I>3p/n, outliers are")
print(outlier2p)

rows_remove2p = which(lev>(2*p/n))
print(rows_remove2p)

developing_LE_no_outliersSTEP2P = developing_LE[-rows_remove2p, ]
head(developing_LE_no_outliersSTEP2P)

nrow(developing_LE)
nrow(developing_LE_no_outliersSTEP2P)
```


2P FIRST MODEL
```{r}
first2plm= lm(Life.expectancy ~  Alcohol + Hepatitis.B  + Measles + BMI + Polio + Diphtheria + HIV.AIDS, data = developing_LE_no_outliersSTEP2P)
summary(first2plm)
#everything signficiant, so no reductions. make int model next

#int model
int2plm = lm(Life.expectancy ~  (Alcohol + Hepatitis.B  + Measles + BMI + Polio + Diphtheria + HIV.AIDS)^2, data = developing_LE_no_outliersSTEP2P)
summary(int2plm)

#reduced int model via stepwise
intred2plm = ols_step_both_p(int2plm, pent = 0.1, prem = 0.3, details = FALSE)
summary(intred2plm$model)


```

WHOA WHOA WHOA STOP HERE DO NOT DO THE 2P MODEL. IT REMOVES A PREDICTOR VARIABLE (MEASLES) WITHOUT REMOVING ITS INTERACTION TERMS. I AM NOT DEALING WITH THAT F THAT.

CAN JUST SAY THAT AFTER CHECKING ASSUMPTIONS FROM THE MODEL WITH OUTLIERS REMOVED OVER THE 3P LEVEL WE DIDNT DO ANY LOWER BECAUSE WE FOUND THAT IT JUST CREATED EVEN WORSE OUTLIERS PRESENT FOR ITS MODEL. SO WE STOPPED THERE AND DIDNT DROP MORE OUTLIERS.




MAKE HIGHER ORDER 2P MODEL
```{r}

#MAKE NEW MODEL HERE AFTER DOING STEPWISE OLS
step_red_2p = lm(Life.expectancy ~  Alcohol + Hepatitis.B  + Measles + BMI + Polio + Diphtheria + HIV.AIDS, data = developing_LE_no_outliersSTEP2P)


#ggpairs(intred2plm)
#make higher order model
higher2plm = lm(Life.expectancy ~ Alcohol + Hepatitis.B + Measles + BMI + Diphtheria + HIV.AIDS + I(HIV.AIDS^2) + I(HIV.AIDS^3) + I(HIV.AIDS^4) + Alcohol:Hepatitis.B + Alcohol:Measles + Alcohol:BMI + Alcohol:Diphtheria + Alcohol:HIV.AIDS + Hepatitis.B:Diphtheria + Measles:Diphtheria + Measles:HIV.AIDS + BMI:Diphtheria + Diphtheria:HIV.AIDS, data = developing_LE_no_outliersSTEP2P)

summary(higher2plm)
bptest(higher2plm)
plot(higher2plm)

shapiro.test(residuals(higher2plm))

#boxcox
bc = boxcox(higher2plm,lambda=seq(-1,1))
bestlambda=bc$x[which(bc$y==max(bc$y))]
bestlambda

```





