---
title: 'Data 603 Term Project'
author: "Ethan Scott, James Huvenaars, Connor Horemans"
date: "2023-11-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data 603 Term Project

```{r}
library(olsrr)
```

## Loading Data

```{r}
df = read.csv('https://raw.githubusercontent.com/ethan2411/Data-603-604/main/603%20Data/500_Cities__Census_Tract-level_Data__GIS_Friendly_Format___2019_release_20231106.csv')
head(df)
```

## Feature Selection

```{r}
#removing all data but the "prevalence" measures for calculations. 
#df.calcs = df frame used for calculations. 
df.calcs <- subset(df, select =- c(StateAbbr, PlaceName, PlaceFIPS, TractFIPS,Place_TractID,Population2010, ACCESS2_Crude95CI, ARTHRITIS_Crude95CI, BINGE_Crude95CI, BPHIGH_Crude95CI, BPMED_Crude95CI, CANCER_Crude95CI, CASTHMA_Crude95CI, CHD_Crude95CI, CHECKUP_Crude95CI, CHOLSCREEN_Crude95CI, COLON_SCREEN_Crude95CI,COPD_Crude95CI, COREM_Crude95CI, COREW_Crude95CI, CSMOKING_Crude95CI, DENTAL_Crude95CI,DIABETES_Crude95CI,  HIGHCHOL_Crude95CI, KIDNEY_Crude95CI, LPA_Crude95CI, MAMMOUSE_Crude95CI, MHLTH_Crude95CI, OBESITY_Crude95CI, PAPTEST_Crude95CI, PHLTH_Crude95CI, SLEEP_Crude95CI, STROKE_Crude95CI, TEETHLOST_Crude95CI, Geolocation))

#head(df.calcs)
```

```{r}
#narrowing the factors to only be unhealthy habits and preventative measures
df.unhealthy.precitors <- df.calcs <- subset(df.calcs, select =- c(ARTHRITIS_CrudePrev, BPHIGH_CrudePrev, CASTHMA_CrudePrev, CHD_CrudePrev, COPD_CrudePrev, DIABETES_CrudePrev, HIGHCHOL_CrudePrev, KIDNEY_CrudePrev, MAMMOUSE_CrudePrev, MHLTH_CrudePrev, PHLTH_CrudePrev, STROKE_CrudePrev))
#head(df.unhealthy.precitors)
```

```{r}
#narrowing the dataset even further to just predictive variables
cancer.pred.only <- subset(df.unhealthy.precitors,  select =- c(BINGE_CrudePrev, CSMOKING_CrudePrev, LPA_CrudePrev, OBESITY_CrudePrev, SLEEP_CrudePrev))
cancer.df = na.omit(data.frame(cancer.pred.only))
head(cancer.df)
```

## Full Linear Model

```{r}
cancer.lm <- lm(CANCER_CrudePrev ~ ., data = cancer.df)
summary(cancer.lm)
```

### Reduced Model

```{r}
#removing ACCESS2_CrudePrev since not statistically significant
cancer.reduced <- lm(CANCER_CrudePrev ~ BPMED_CrudePrev + CHECKUP_CrudePrev + CHOLSCREEN_CrudePrev + COLON_SCREEN_CrudePrev + COREM_CrudePrev + COREW_CrudePrev  + DENTAL_CrudePrev + PAPTEST_CrudePrev + TEETHLOST_CrudePrev, data = cancer.df)

summary(cancer.reduced)

#p-value from anova is greater than 0.05, reduced model is not significantly different than full model
anova(cancer.lm, cancer.reduced)
```

According to ANOVA, we are good to drop ACCESS2_CrudePrev.

```{r echo=TRUE}
#testing to see if this meets the assumptions of the model.
par(mfrow = c(2, 2))  # 2 rows, 2 columns
plot(cancer.reduced, which = 1)
plot(cancer.reduced, which = 2)
plot(cancer.reduced, which = 3)
plot(cancer.reduced, which = 4)
```

The model does NOT meet the criteria of the model.

```{r}
pairs(cancer.df)
```

Given the ANOVA output, we can drop ACCESS2_CrudePrev.

#### Box Cox Transformation on Reduced

```{r}
library(car)
bc = boxCox(cancer.reduced,lambda=seq(-1,1))

bestlambda=bc$x[which(bc$y==max(bc$y))]
bestlambda
```

```{r}
library(dplyr)
cancer.boxcox = cancer.df %>%
  mutate(CANCER_boxcox = (CANCER_CrudePrev^bestlambda - 1)/bestlambda)

head(cancer.boxcox)
```

```{r}
bc.cancer <- lm(CANCER_boxcox ~ BPMED_CrudePrev + CHECKUP_CrudePrev + CHOLSCREEN_CrudePrev + COLON_SCREEN_CrudePrev + COREM_CrudePrev + COREW_CrudePrev  + DENTAL_CrudePrev + PAPTEST_CrudePrev + TEETHLOST_CrudePrev, data = cancer.boxcox)
summary(bc.cancer)
```

```{r}
#still not good
plot(bc.cancer)
```

#### Log Transformation on Reduced

```{r}
# Log transform the response variable
cancer.reduced_log <- lm(log(CANCER_CrudePrev) ~ log(BPMED_CrudePrev) + log(CHECKUP_CrudePrev) + log(CHOLSCREEN_CrudePrev) + log(COLON_SCREEN_CrudePrev) + log(COREM_CrudePrev) + log(COREW_CrudePrev) + log(DENTAL_CrudePrev) + log(PAPTEST_CrudePrev) + log(TEETHLOST_CrudePrev), data = cancer.df)
summary(cancer.reduced_log)
#Variances are ew, gross, yucky...
#normality has smaller lower tail, upper tail is gross though
plot(cancer.reduced_log)
```

```{r}
cancer.log = lm(log(CANCER_CrudePrev) ~ BPMED_CrudePrev + CHECKUP_CrudePrev + CHOLSCREEN_CrudePrev + COLON_SCREEN_CrudePrev + COREM_CrudePrev + COREW_CrudePrev  + DENTAL_CrudePrev + PAPTEST_CrudePrev + TEETHLOST_CrudePrev, data = cancer.df)
plot(cancer.log)
```

#### Square Root Transformation on Reduced

```{r}
# Square root transform the response variable
cancer.reduced_sqrt <- lm(sqrt(CANCER_CrudePrev) ~ sqrt(BPMED_CrudePrev) + sqrt(CHECKUP_CrudePrev) + sqrt(CHOLSCREEN_CrudePrev) + sqrt(COLON_SCREEN_CrudePrev) + sqrt(COREM_CrudePrev) + sqrt(COREW_CrudePrev) + sqrt(DENTAL_CrudePrev) + sqrt(PAPTEST_CrudePrev) + sqrt(TEETHLOST_CrudePrev), data = cancer.df)
summary(cancer.reduced_sqrt)
#this one also needs jail
plot(cancer.reduced_sqrt)
```

```{r}
cancer.sqrt = lm(sqrt(CANCER_CrudePrev) ~ BPMED_CrudePrev + CHECKUP_CrudePrev + CHOLSCREEN_CrudePrev + COLON_SCREEN_CrudePrev + COREM_CrudePrev + COREW_CrudePrev  + DENTAL_CrudePrev + PAPTEST_CrudePrev + TEETHLOST_CrudePrev, data = cancer.df)
plot(cancer.sqrt)
```

#### Inverse Transformation

```{r}
cancer.reduced_inverse <- lm(I(1/CANCER_CrudePrev) ~ I(1/BPMED_CrudePrev) + I(1/CHECKUP_CrudePrev) + I(1/CHOLSCREEN_CrudePrev) + I(1/COLON_SCREEN_CrudePrev) + I(1/COREM_CrudePrev) + I(1/COREW_CrudePrev) + I(1/DENTAL_CrudePrev) + I(1/PAPTEST_CrudePrev) + I(1/TEETHLOST_CrudePrev), data = cancer.df)
summary(cancer.reduced_inverse)
#what the actual fuck is this?
plot(cancer.reduced_inverse)
```

```{r}
cancer.inv = lm(I(1/CANCER_CrudePrev) ~ BPMED_CrudePrev + CHECKUP_CrudePrev + CHOLSCREEN_CrudePrev + COLON_SCREEN_CrudePrev + COREM_CrudePrev + COREW_CrudePrev  + DENTAL_CrudePrev + PAPTEST_CrudePrev + TEETHLOST_CrudePrev, data = cancer.df)
plot(cancer.inv)
```

## Interactions Model

```{r}
cancer.interactions <- lm(CANCER_CrudePrev ~ (BPMED_CrudePrev + CHECKUP_CrudePrev + CHOLSCREEN_CrudePrev + COLON_SCREEN_CrudePrev + COREM_CrudePrev + COREW_CrudePrev  + DENTAL_CrudePrev + PAPTEST_CrudePrev + TEETHLOST_CrudePrev)^2, data = cancer.df)

summary(cancer.interactions)
```

## Stepwise Model (AIC Prioritized)

```{r}
model_step <- step(cancer.interactions, direction = "both")
```

```{r}
summary(model_step)
```

```{r}
rounded_coefficients <- round(coefficients(model_step), 6)  # Round coefficients to 6 digits

# Construct the formula with rounded coefficients
formula_str <- paste0("y ~ ", 
                     round(rounded_coefficients[1], 6), 
                     "", 
                     paste(sprintf(" %+.6f*%s ", 
                                   rounded_coefficients[-1], 
                                   names(rounded_coefficients[-1])), 
                           collapse="")
                     )

# Convert the formula string into a formula object
formula_with_rounded_coefficients <- as.formula(formula_str)

# Print the formula
print(formula_with_rounded_coefficients)
```

## Model Assumptions

```{r}
#linearity assumption
library(ggplot2)
ggplot(model_step, aes(x=.fitted, y=.resid)) +
geom_point() + geom_smooth()+
geom_hline(yintercept = 0)
```

```{r}
plot(model_step)
```

### Box Cox Transformation

```{r}
library(MASS)
bc = boxcox(model_step,lambda=seq(-1,1))

bestlambda=bc$x[which(bc$y==max(bc$y))]
bestlambda
```

```{r}
library(mosaic)
cancer.pred.boxcox = cancer.df %>%
  mutate(CANCER_boxcox = (CANCER_CrudePrev^bestlambda - 1)/bestlambda)

head(cancer.pred.boxcox)
```

#### Interactions with box cox

```{r}
interactions.boxcox = lm(formula = CANCER_boxcox ~ BPMED_CrudePrev + CHECKUP_CrudePrev + 
    CHOLSCREEN_CrudePrev + COLON_SCREEN_CrudePrev + COREM_CrudePrev + 
    COREW_CrudePrev + DENTAL_CrudePrev + PAPTEST_CrudePrev + 
    TEETHLOST_CrudePrev + BPMED_CrudePrev:CHECKUP_CrudePrev + 
    BPMED_CrudePrev:CHOLSCREEN_CrudePrev + BPMED_CrudePrev:COLON_SCREEN_CrudePrev + 
    BPMED_CrudePrev:COREM_CrudePrev + BPMED_CrudePrev:COREW_CrudePrev + 
    BPMED_CrudePrev:PAPTEST_CrudePrev + BPMED_CrudePrev:TEETHLOST_CrudePrev + 
    CHECKUP_CrudePrev:CHOLSCREEN_CrudePrev + CHECKUP_CrudePrev:COLON_SCREEN_CrudePrev + 
    CHECKUP_CrudePrev:COREM_CrudePrev + CHECKUP_CrudePrev:DENTAL_CrudePrev + 
    CHECKUP_CrudePrev:PAPTEST_CrudePrev + CHECKUP_CrudePrev:TEETHLOST_CrudePrev + 
    CHOLSCREEN_CrudePrev:COLON_SCREEN_CrudePrev + CHOLSCREEN_CrudePrev:COREM_CrudePrev + 
    CHOLSCREEN_CrudePrev:COREW_CrudePrev + CHOLSCREEN_CrudePrev:DENTAL_CrudePrev + 
    CHOLSCREEN_CrudePrev:PAPTEST_CrudePrev + CHOLSCREEN_CrudePrev:TEETHLOST_CrudePrev + 
    COLON_SCREEN_CrudePrev:COREW_CrudePrev + COLON_SCREEN_CrudePrev:DENTAL_CrudePrev + 
    COLON_SCREEN_CrudePrev:PAPTEST_CrudePrev + COLON_SCREEN_CrudePrev:TEETHLOST_CrudePrev + 
    COREM_CrudePrev:COREW_CrudePrev + COREM_CrudePrev:DENTAL_CrudePrev + 
    COREM_CrudePrev:PAPTEST_CrudePrev + COREW_CrudePrev:DENTAL_CrudePrev + 
    COREW_CrudePrev:PAPTEST_CrudePrev + COREW_CrudePrev:TEETHLOST_CrudePrev + 
    DENTAL_CrudePrev:PAPTEST_CrudePrev + DENTAL_CrudePrev:TEETHLOST_CrudePrev + 
    PAPTEST_CrudePrev:TEETHLOST_CrudePrev, data = cancer.pred.boxcox)

plot(interactions.boxcox)
summary(interactions.boxcox)
```

```{r}
summary(interactions.boxcox)$sigma

```

```{r}
library(lmtest)
bptest(interactions.boxcox)
```

Based on the above low p-value, we can assume there is heteroscedasticity in the model. We will need further trasnformations to the model to meet the needed assumption.

```{r}
#shapiro.test(residuals(interactions.boxcox))
#Error in shapiro.test(residuals(interactions.boxcox)) :sample size must be between 3 and 5000
```

Since we cannot complete the Shapiro Wilk Test because our sample size is too large, we must rely on qqplot to test for normality.

```{r}
#perform Kolmogorov-Smirnov test (2 sample test, creating a new sample that is normal)
ks.test(x = residuals( interactions.boxcox), y = rnorm(mean = 0, sd = sqrt(0.08362472), 24825))
```

Based on the KS test, the distribution is not normal.

## Stepwise Model Prioritizing BIC instead of AIC

```{r}
model_step_bic <- step(cancer.interactions, direction = "both", k = log(24826))

```

```{r}
summary(model_step_bic)
```

Getting a nice formula to display.

```{r}
rounded_coefficients <- round(coefficients(model_step_bic), 6)  # Round coefficients to 6 digits

# Construct the formula with rounded coefficients
formula_str <- paste0("y ~ ", 
                     round(rounded_coefficients[1], 6), 
                     "", 
                     paste(sprintf(" %+.6f*%s ", 
                                   rounded_coefficients[-1], 
                                   names(rounded_coefficients[-1])), 
                           collapse="")
                     )

# Convert the formula string into a formula object
formula_with_rounded_coefficients <- as.formula(formula_str)

# Print the formula
print(formula_with_rounded_coefficients)
```

```{r}
plot(model_step_bic)
```

```{r}
bc.bic = boxcox(model_step_bic,lambda=seq(-1,1))

bestlambda.bic=bc.bic$x[which(bc.bic$y==max(bc.bic$y))]
bestlambda.bic
```

```{r}
cancer.pred.boxcox = cancer.df %>%
  mutate(CANCER_boxcox_bic = (CANCER_CrudePrev^bestlambda.bic - 1)/bestlambda.bic)

head(cancer.pred.boxcox)
```

```{r}
interactions.boxcox.bic <- lm(CANCER_boxcox_bic ~ BPMED_CrudePrev + CHECKUP_CrudePrev + 
    CHOLSCREEN_CrudePrev + COLON_SCREEN_CrudePrev + COREM_CrudePrev + 
    COREW_CrudePrev + DENTAL_CrudePrev + PAPTEST_CrudePrev + 
    TEETHLOST_CrudePrev + BPMED_CrudePrev:CHECKUP_CrudePrev + 
    BPMED_CrudePrev:CHOLSCREEN_CrudePrev + BPMED_CrudePrev:COLON_SCREEN_CrudePrev + 
    BPMED_CrudePrev:COREM_CrudePrev + BPMED_CrudePrev:COREW_CrudePrev + 
    BPMED_CrudePrev:PAPTEST_CrudePrev + BPMED_CrudePrev:TEETHLOST_CrudePrev + 
    CHECKUP_CrudePrev:CHOLSCREEN_CrudePrev + CHECKUP_CrudePrev:COLON_SCREEN_CrudePrev + 
    CHECKUP_CrudePrev:COREM_CrudePrev + CHECKUP_CrudePrev:DENTAL_CrudePrev + 
    CHECKUP_CrudePrev:PAPTEST_CrudePrev + CHECKUP_CrudePrev:TEETHLOST_CrudePrev + 
    CHOLSCREEN_CrudePrev:COLON_SCREEN_CrudePrev + CHOLSCREEN_CrudePrev:COREM_CrudePrev + 
    CHOLSCREEN_CrudePrev:COREW_CrudePrev + CHOLSCREEN_CrudePrev:DENTAL_CrudePrev + 
    CHOLSCREEN_CrudePrev:PAPTEST_CrudePrev + CHOLSCREEN_CrudePrev:TEETHLOST_CrudePrev + 
    COLON_SCREEN_CrudePrev:COREW_CrudePrev + COLON_SCREEN_CrudePrev:DENTAL_CrudePrev + 
    COLON_SCREEN_CrudePrev:PAPTEST_CrudePrev + COLON_SCREEN_CrudePrev:TEETHLOST_CrudePrev + 
    COREM_CrudePrev:COREW_CrudePrev + COREM_CrudePrev:DENTAL_CrudePrev + 
    COREM_CrudePrev:PAPTEST_CrudePrev + COREW_CrudePrev:DENTAL_CrudePrev + 
    COREW_CrudePrev:PAPTEST_CrudePrev + COREW_CrudePrev:TEETHLOST_CrudePrev + 
    DENTAL_CrudePrev:PAPTEST_CrudePrev + DENTAL_CrudePrev:TEETHLOST_CrudePrev + 
    PAPTEST_CrudePrev:TEETHLOST_CrudePrev, data = cancer.pred.boxcox)

plot(interactions.boxcox.bic)
```

### Additive model removing highest VIF value (DENTAL_CrudePrev)

Since we did not satify the assumption of homoscedasticity, return to VIF to remove highest VIF value (DENTAL_CrudePrev)

```{r}
library(car)
vif(lm(CANCER_CrudePrev ~ BPMED_CrudePrev + CHECKUP_CrudePrev + CHOLSCREEN_CrudePrev + COLON_SCREEN_CrudePrev + COREM_CrudePrev + COREW_CrudePrev  + DENTAL_CrudePrev + PAPTEST_CrudePrev + TEETHLOST_CrudePrev, data = df.unhealthy.precitors))


```

```{r}
#removing DENTAL_CrudePrev since it has the highest VIF value

additive.nodental <- lm(CANCER_CrudePrev ~ BPMED_CrudePrev + CHECKUP_CrudePrev + CHOLSCREEN_CrudePrev + COLON_SCREEN_CrudePrev + COREM_CrudePrev + COREW_CrudePrev + PAPTEST_CrudePrev + TEETHLOST_CrudePrev, data = df.unhealthy.precitors)

summary(additive.nodental)

anova(cancer.lm, additive.nodental)
```

We cannot remove DENTAL_CrudePrev since the ANOVA value is less than 0.05. Try removing the variable with the second highest VIF.

### Additive model removing second highest VIF value (TEETHLOST_CrudePrev)

```{r}
#removing TEETHLOST_CrudePrev since it has the second highest VIF value, and we cannot remove the term with the highest VIF according to the ANOVA test. 

additive.noteethlost <- lm(CANCER_CrudePrev ~ BPMED_CrudePrev + CHECKUP_CrudePrev + CHOLSCREEN_CrudePrev + COLON_SCREEN_CrudePrev + COREM_CrudePrev + COREW_CrudePrev  + DENTAL_CrudePrev + PAPTEST_CrudePrev, data = df.unhealthy.precitors)

summary(additive.noteethlost)

anova(cancer.lm, additive.noteethlost)


```

Since the P-value of the ANOVA is less than 0.05, we cannnot drop TEETHLOST_CrudePrev.

```{r}
?step
```

## Transformations on Interactions(Using AIC Model)

### Log

```{r}
interactions.log = lm(log(CANCER_CrudePrev) ~ log(BPMED_CrudePrev) + log(CHECKUP_CrudePrev) + log(CHOLSCREEN_CrudePrev) + log(COLON_SCREEN_CrudePrev) + log(COREM_CrudePrev) + log(COREW_CrudePrev) + log(DENTAL_CrudePrev) + log(PAPTEST_CrudePrev) + log(TEETHLOST_CrudePrev) + log(BPMED_CrudePrev * CHECKUP_CrudePrev) + log(BPMED_CrudePrev * CHOLSCREEN_CrudePrev) + log(BPMED_CrudePrev * COLON_SCREEN_CrudePrev) + log(BPMED_CrudePrev * COREM_CrudePrev) + log(BPMED_CrudePrev * COREW_CrudePrev) + log(BPMED_CrudePrev * PAPTEST_CrudePrev) + log(BPMED_CrudePrev * TEETHLOST_CrudePrev) + log(CHECKUP_CrudePrev * CHOLSCREEN_CrudePrev) + log(CHECKUP_CrudePrev * COLON_SCREEN_CrudePrev) + log(CHECKUP_CrudePrev * COREM_CrudePrev) + log(CHECKUP_CrudePrev * DENTAL_CrudePrev) + log(CHECKUP_CrudePrev * PAPTEST_CrudePrev) + log(CHECKUP_CrudePrev * TEETHLOST_CrudePrev) + log(CHOLSCREEN_CrudePrev * COLON_SCREEN_CrudePrev) + log(CHOLSCREEN_CrudePrev * COREM_CrudePrev) + log(CHOLSCREEN_CrudePrev * COREW_CrudePrev) + log(CHOLSCREEN_CrudePrev * DENTAL_CrudePrev) + log(CHOLSCREEN_CrudePrev * PAPTEST_CrudePrev) + log(CHOLSCREEN_CrudePrev * TEETHLOST_CrudePrev) + log(COLON_SCREEN_CrudePrev * COREW_CrudePrev) + log(COLON_SCREEN_CrudePrev * DENTAL_CrudePrev) + log(COLON_SCREEN_CrudePrev * PAPTEST_CrudePrev) + log(COLON_SCREEN_CrudePrev * TEETHLOST_CrudePrev) + log(COREM_CrudePrev * COREW_CrudePrev) + log(COREM_CrudePrev * DENTAL_CrudePrev) + log(COREM_CrudePrev * PAPTEST_CrudePrev) + log(COREW_CrudePrev * DENTAL_CrudePrev) + log(COREW_CrudePrev * PAPTEST_CrudePrev) + log(COREW_CrudePrev * TEETHLOST_CrudePrev) + log(DENTAL_CrudePrev * PAPTEST_CrudePrev) + log(DENTAL_CrudePrev * TEETHLOST_CrudePrev) + log(PAPTEST_CrudePrev * TEETHLOST_CrudePrev), data = cancer.df)

```

```{r}
plot(interactions.log)
```

### SQRT

```{r}
interactions.sqrt = lm(sqrt(CANCER_CrudePrev) ~ sqrt(BPMED_CrudePrev) + sqrt(CHECKUP_CrudePrev) + sqrt(CHOLSCREEN_CrudePrev) + sqrt(COLON_SCREEN_CrudePrev) + sqrt(COREM_CrudePrev) + sqrt(COREW_CrudePrev) + sqrt(DENTAL_CrudePrev) + sqrt(PAPTEST_CrudePrev) + sqrt(TEETHLOST_CrudePrev) + sqrt(BPMED_CrudePrev * CHECKUP_CrudePrev) + sqrt(BPMED_CrudePrev * CHOLSCREEN_CrudePrev) + sqrt(BPMED_CrudePrev * COLON_SCREEN_CrudePrev) + sqrt(BPMED_CrudePrev * COREM_CrudePrev) + sqrt(BPMED_CrudePrev * COREW_CrudePrev) + sqrt(BPMED_CrudePrev * PAPTEST_CrudePrev) + sqrt(BPMED_CrudePrev * TEETHLOST_CrudePrev) + sqrt(CHECKUP_CrudePrev * CHOLSCREEN_CrudePrev) + sqrt(CHECKUP_CrudePrev * COLON_SCREEN_CrudePrev) + sqrt(CHECKUP_CrudePrev * COREM_CrudePrev) + sqrt(CHECKUP_CrudePrev * DENTAL_CrudePrev) + sqrt(CHECKUP_CrudePrev * PAPTEST_CrudePrev) + sqrt(CHECKUP_CrudePrev * TEETHLOST_CrudePrev) + sqrt(CHOLSCREEN_CrudePrev * COLON_SCREEN_CrudePrev) + sqrt(CHOLSCREEN_CrudePrev * COREM_CrudePrev) + sqrt(CHOLSCREEN_CrudePrev * COREW_CrudePrev) + sqrt(CHOLSCREEN_CrudePrev * DENTAL_CrudePrev) + sqrt(CHOLSCREEN_CrudePrev * PAPTEST_CrudePrev) + sqrt(CHOLSCREEN_CrudePrev * TEETHLOST_CrudePrev) + sqrt(COLON_SCREEN_CrudePrev * COREW_CrudePrev) + sqrt(COLON_SCREEN_CrudePrev * DENTAL_CrudePrev) + sqrt(COLON_SCREEN_CrudePrev * PAPTEST_CrudePrev) + sqrt(COLON_SCREEN_CrudePrev * TEETHLOST_CrudePrev) + sqrt(COREM_CrudePrev * COREW_CrudePrev) + sqrt(COREM_CrudePrev * DENTAL_CrudePrev) + sqrt(COREM_CrudePrev * PAPTEST_CrudePrev) + sqrt(COREW_CrudePrev * DENTAL_CrudePrev) + sqrt(COREW_CrudePrev * PAPTEST_CrudePrev) + sqrt(COREW_CrudePrev * TEETHLOST_CrudePrev) + sqrt(DENTAL_CrudePrev * PAPTEST_CrudePrev) + sqrt(DENTAL_CrudePrev * TEETHLOST_CrudePrev) + sqrt(PAPTEST_CrudePrev * TEETHLOST_CrudePrev), data = cancer.df)

```

```{r}
plot(interactions.sqrt)
```

### Inverse

```{r}
interactions.inv = lm(I(1/CANCER_CrudePrev) ~ I(1/BPMED_CrudePrev) + I(1/CHECKUP_CrudePrev) + I(1/CHOLSCREEN_CrudePrev) + I(1/COLON_SCREEN_CrudePrev) + I(1/COREM_CrudePrev) + I(1/COREW_CrudePrev) + I(1/DENTAL_CrudePrev) + I(1/PAPTEST_CrudePrev) + I(1/TEETHLOST_CrudePrev) + I(1/(BPMED_CrudePrev * CHECKUP_CrudePrev)) + I(1/(BPMED_CrudePrev * CHOLSCREEN_CrudePrev)) + I(1/(BPMED_CrudePrev * COLON_SCREEN_CrudePrev)) + I(1/(BPMED_CrudePrev * COREM_CrudePrev)) + I(1/(BPMED_CrudePrev * COREW_CrudePrev)) + I(1/(BPMED_CrudePrev * PAPTEST_CrudePrev)) + I(1/(BPMED_CrudePrev * TEETHLOST_CrudePrev)) + I(1/(CHECKUP_CrudePrev * CHOLSCREEN_CrudePrev)) + I(1/(CHECKUP_CrudePrev * COLON_SCREEN_CrudePrev)) + I(1/(CHECKUP_CrudePrev * COREM_CrudePrev)) + I(1/(CHECKUP_CrudePrev * DENTAL_CrudePrev)) + I(1/(CHECKUP_CrudePrev * PAPTEST_CrudePrev)) + I(1/(CHECKUP_CrudePrev * TEETHLOST_CrudePrev)) + I(1/(CHOLSCREEN_CrudePrev * COLON_SCREEN_CrudePrev)) + I(1/(CHOLSCREEN_CrudePrev * COREM_CrudePrev)) + I(1/(CHOLSCREEN_CrudePrev * COREW_CrudePrev)) + I(1/(CHOLSCREEN_CrudePrev * DENTAL_CrudePrev)) + I(1/(CHOLSCREEN_CrudePrev * PAPTEST_CrudePrev)) + I(1/(CHOLSCREEN_CrudePrev * TEETHLOST_CrudePrev)) + I(1/(COLON_SCREEN_CrudePrev * COREW_CrudePrev)) + I(1/(COLON_SCREEN_CrudePrev * DENTAL_CrudePrev)) + I(1/(COLON_SCREEN_CrudePrev * PAPTEST_CrudePrev)) + I(1/(COLON_SCREEN_CrudePrev * TEETHLOST_CrudePrev)) + I(1/(COREM_CrudePrev * COREW_CrudePrev)) + I(1/(COREM_CrudePrev * DENTAL_CrudePrev)) + I(1/(COREM_CrudePrev * PAPTEST_CrudePrev)) + I(1/(COREW_CrudePrev * DENTAL_CrudePrev)) + I(1/(COREW_CrudePrev * PAPTEST_CrudePrev)) + I(1/(COREW_CrudePrev * TEETHLOST_CrudePrev)) + I(1/(DENTAL_CrudePrev * PAPTEST_CrudePrev)) + I(1/(DENTAL_CrudePrev * TEETHLOST_CrudePrev)) + I(1/(PAPTEST_CrudePrev * TEETHLOST_CrudePrev)), data = cancer.df)

```

```{r}
plot(interactions.inv)
```

## Polynomial

```{r}
polynomial.df = na.omit(cancer.df)
head(polynomial.df)
```

```{r}
# Quadratic Polynomial Regression
cancer.quadratic <- lm(CANCER_CrudePrev ~ poly(BPMED_CrudePrev, 2) + poly(CHECKUP_CrudePrev, 2) + poly(CHOLSCREEN_CrudePrev, 2) + poly(COLON_SCREEN_CrudePrev, 2) + poly(COREM_CrudePrev, 2) + poly(COREW_CrudePrev, 2) + poly(DENTAL_CrudePrev, 2) + poly(PAPTEST_CrudePrev, 2) + poly(TEETHLOST_CrudePrev, 2), data = polynomial.df)
#all significant so thats promissing
summary(cancer.quadratic)
#gross
plot(cancer.quadratic)
```

```{r}
# 3rd Degree Polynomial Regression
cancer.cubic <- lm(CANCER_CrudePrev ~ poly(BPMED_CrudePrev, 3) + poly(CHECKUP_CrudePrev, 3) + poly(CHOLSCREEN_CrudePrev, 3) + poly(COLON_SCREEN_CrudePrev, 3) + poly(COREM_CrudePrev, 3) + poly(COREW_CrudePrev, 3) + poly(DENTAL_CrudePrev, 3) + poly(PAPTEST_CrudePrev, 3) + poly(TEETHLOST_CrudePrev, 3), data = polynomial.df)
#still not bad
summary(cancer.cubic)
#ew yuck
plot(cancer.cubic)
```

```{r}
# 4th Degree Polynomial Regression
cancer.quartic <- lm(CANCER_CrudePrev ~ poly(BPMED_CrudePrev, 4) + poly(CHECKUP_CrudePrev, 4) + poly(CHOLSCREEN_CrudePrev, 4) + poly(COLON_SCREEN_CrudePrev, 4) + poly(COREM_CrudePrev, 4) + poly(COREW_CrudePrev, 4) + poly(DENTAL_CrudePrev, 4) + poly(PAPTEST_CrudePrev, 4) + poly(TEETHLOST_CrudePrev, 4), data = polynomial.df)

summary(cancer.quartic)
#this is actually looking more prommising.... soooooo im going ham next
plot(cancer.quartic)
```

```{r}
# 8th Degree Polynomial Regression
cancer.8 <- lm(CANCER_CrudePrev ~ poly(BPMED_CrudePrev, 8) + poly(CHECKUP_CrudePrev, 8) + poly(CHOLSCREEN_CrudePrev, 8) + poly(COLON_SCREEN_CrudePrev, 8) + poly(COREM_CrudePrev, 8) + poly(COREW_CrudePrev, 8) + poly(DENTAL_CrudePrev, 8) + poly(PAPTEST_CrudePrev, 8) + poly(TEETHLOST_CrudePrev, 8), data = polynomial.df)

summary(cancer.8)
#yucky
plot(cancer.8)
```

## LASSO on Polynomial Regression (Not looking good)

```{r}
library(glmnet)
```

```{r}
#Getting train and test splits
y_train = cancer.df$CANCER_CrudePrev
X_train = model.matrix(CANCER_CrudePrev ~ poly(BPMED_CrudePrev, 8) + poly(CHECKUP_CrudePrev, 8) + poly(CHOLSCREEN_CrudePrev, 8) + poly(COLON_SCREEN_CrudePrev, 8) + poly(COREM_CrudePrev, 8) + poly(COREW_CrudePrev, 8) + poly(DENTAL_CrudePrev, 8) + poly(PAPTEST_CrudePrev, 8) + poly(TEETHLOST_CrudePrev, 8), data = cancer.df)[, -3]#3rd column is CANCER

# Perform Lasso regression with cross-validation to select the lambda parameter
cv.lasso = cv.glmnet(X_train, y_train, alpha = 1)

# Fit the final Lasso model with the selected lambda using the training data
lasso_model = glmnet(X_train, y_train, alpha = 1, lambda = cv.lasso$lambda.min)

#THIS ISNT RIGHT, SHOULD HAVE TRAIN AND TEST SPLITS BUT JUST USING WHOLE DATA FOR RN
predictions = predict(lasso_model, s = cv.lasso$lambda.min, newx = X_train)
```

```{r}
coefficients(lasso_model)
```

```{r}
# Calculate residuals
residuals = y_train-predictions
# Normality of Residuals
# Q-Q Plot
qqnorm(residuals)
qqline(residuals)
```

```{r}
plot(predictions, residuals, main = "Residuals vs. Fitted Values", xlab = "Fitted Values", ylab = "Residuals")
abline(h = 0, col = "red")
```
