---
title: 'Data 603 Term Project'
author: "Ethan Scott, James Huvenaars, Connor Horemans"
date: "2023-11-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(mctest)
library(car)
library(ggplot2)
library(GGally)
library(MASS)
library(mosaic)
library(lmtest)
library(olsrr)
```

# Data 603 Term Project

```{r}
df = read.csv('https://raw.githubusercontent.com/ethan2411/Data-603-604/main/603%20Data/500_Cities__Census_Tract-level_Data__GIS_Friendly_Format___2019_release_20231106.csv')
head(df)
```

```{r}
#removing all data but the Cancer crude prevalence and preventative measures. 

#df.calcs = df frame used for calculations. 
df.calcs <- subset(df, select = c(CANCER_CrudePrev, BINGE_CrudePrev, CSMOKING_CrudePrev, LPA_CrudePrev, OBESITY_CrudePrev, SLEEP_CrudePrev))

head(df.calcs)
```

```{r}
additive <- lm(CANCER_CrudePrev ~ ., df.calcs)
summary(additive)
```
```{r}
imcdiag(additive, method = 'VIF')
```

```{r}
vif(additive)
```


```{r}
additive.noLPA <- lm(CANCER_CrudePrev ~ BINGE_CrudePrev + CSMOKING_CrudePrev + OBESITY_CrudePrev + SLEEP_CrudePrev, data = df.calcs)
vif(additive.noLPA)
# all now have VIFS off Less than 5. let's take it from the top. 
```



```{r}
additive.interactions <- lm(CANCER_CrudePrev ~ (.)^2, df.calcs)
summary(additive.interactions)
```
```{r}
plot(additive.interactions)
bptest(additive.interactions)
```

#log transformation

```{r}
additive.log <- lm(log(CANCER_CrudePrev) ~ (.)^2, df.calcs)
summary(additive.log)
plot(additive.log)
bptest(additive.log)
```

Log sucks balls, try boxcox


```{r}
#remove LPA_CrudePrev:SLEEP_CrudePrev since it has high p-value. 
additive.log2 <- lm(log(CANCER_CrudePrev) ~ BINGE_CrudePrev + CSMOKING_CrudePrev + LPA_CrudePrev + OBESITY_CrudePrev + SLEEP_CrudePrev + BINGE_CrudePrev:LPA_CrudePrev + BINGE_CrudePrev:OBESITY_CrudePrev + BINGE_CrudePrev:SLEEP_CrudePrev + CSMOKING_CrudePrev:LPA_CrudePrev + CSMOKING_CrudePrev:OBESITY_CrudePrev + CSMOKING_CrudePrev:SLEEP_CrudePrev + LPA_CrudePrev:OBESITY_CrudePrev + OBESITY_CrudePrev:SLEEP_CrudePrev + BINGE_CrudePrev:CSMOKING_CrudePrev , df.calcs)

summary(additive.log2)
plot(additive.log2)
bptest(additive.log2)
anova(additive.log, additive.log2)
````

```{r}
#remove BINGE_CrudePrev:CSMOKING_CrudePrev since it has high p-value. 
additive.log3 <- lm(log(CANCER_CrudePrev) ~ BINGE_CrudePrev + CSMOKING_CrudePrev + LPA_CrudePrev + OBESITY_CrudePrev + SLEEP_CrudePrev + BINGE_CrudePrev:LPA_CrudePrev + BINGE_CrudePrev:OBESITY_CrudePrev + BINGE_CrudePrev:SLEEP_CrudePrev + CSMOKING_CrudePrev:LPA_CrudePrev + CSMOKING_CrudePrev:OBESITY_CrudePrev + CSMOKING_CrudePrev:SLEEP_CrudePrev + LPA_CrudePrev:OBESITY_CrudePrev + OBESITY_CrudePrev:SLEEP_CrudePrev, df.calcs)

summary(additive.log3)
plot(additive.log3)
bptest(additive.log3)
anova(additive.log, additive.log3)
````

```{r}
#remove SMOKING_CrudePrev:SLEEP_CrudePrev since it has high p-value. 
additive.log4 <- lm(log(CANCER_CrudePrev) ~ BINGE_CrudePrev + CSMOKING_CrudePrev + LPA_CrudePrev + OBESITY_CrudePrev + SLEEP_CrudePrev + BINGE_CrudePrev:LPA_CrudePrev + BINGE_CrudePrev:OBESITY_CrudePrev + BINGE_CrudePrev:SLEEP_CrudePrev + CSMOKING_CrudePrev:LPA_CrudePrev + CSMOKING_CrudePrev:OBESITY_CrudePrev + LPA_CrudePrev:OBESITY_CrudePrev + OBESITY_CrudePrev:SLEEP_CrudePrev, df.calcs)

summary(additive.log4)
plot(additive.log4)
bptest(additive.log4)
anova(additive.log3, additive.log4)
````


```{r}
#remove BINGE_CrudePrev:OBESITY_CrudePrev since it has high p-value. 
additive.log5 <- lm(log(CANCER_CrudePrev) ~ BINGE_CrudePrev + CSMOKING_CrudePrev + LPA_CrudePrev + OBESITY_CrudePrev + SLEEP_CrudePrev + BINGE_CrudePrev:LPA_CrudePrev + BINGE_CrudePrev:SLEEP_CrudePrev + CSMOKING_CrudePrev:LPA_CrudePrev + CSMOKING_CrudePrev:OBESITY_CrudePrev + LPA_CrudePrev:OBESITY_CrudePrev + OBESITY_CrudePrev:SLEEP_CrudePrev, df.calcs)

summary(additive.log5)
plot(additive.log5, which = 5)
bptest(additive.log5)
anova(additive.log5, additive.log4)
````
```{r}
df.calcs[cooks.distance(additive.log5)>0.5,]
plot(additive.log5,pch=18,col="red",which=c(4)) #which =4 only prints the cook distance plot. 
```


#boxcox transformation

```{r}
bc = boxcox(additive.interactions,lambda=seq(-1,1))

bestlambda=bc$x[which(bc$y==max(bc$y))]
bestlambda
```
```{r}
bcmodel=lm((((CANCER_CrudePrev^bestlambda)-1)/bestlambda) ~ (.)^2, df.calcs)
summary(bcmodel)
```
```{r}
plot(bcmodel)
bptest(bcmodel)
```

# square root transformation

```{r}
additive.sqrt <- lm(sqrt(CANCER_CrudePrev) ~ (.)^2, df.calcs)
summary(additive.sqrt)
plot(additive.sqrt)
bptest(additive.sqrt)
```

#Checking for square root transformations: 

```{r}
ggpairs(df.calcs, lower = list(continuous = "smooth_loess", combo = "facethist", discrete = "facetbar", na = "na"))

```

#CANCER AND LPA SEEM A BIT ODD, try to fix that with some higher powers:


```{r}
plot(additive)
```
```{r}
additive.squared <- lm(CANCER_CrudePrev ~ BINGE_CrudePrev + CSMOKING_CrudePrev + OBESITY_CrudePrev + SLEEP_CrudePrev + LPA_CrudePrev + I(LPA_CrudePrev^2), data = df.calcs)
summary(additive.squared)
plot(additive.squared, which = 1)
```


```{r}
more.power <- lm(CANCER_CrudePrev ~ poly(LPA_CrudePrev,11,raw = T) + BINGE_CrudePrev + CSMOKING_CrudePrev + OBESITY_CrudePrev + SLEEP_CrudePrev + LPA_CrudePrev, data = df.calcs)
summary(more.power)
plot(more.power, which = 1)
bptest(more.power)
```

this sucks too. 




#Remove the highest VIF value (LPA and see how that does)

```{r}
additive.noLPA <- lm(CANCER_CrudePrev ~ BINGE_CrudePrev + CSMOKING_CrudePrev + OBESITY_CrudePrev + SLEEP_CrudePrev, data = df.calcs)
vif(additive.noLPA)
# all now have VIFS off Less than 5. let's take it from the top. 
```

##Additive Model
```{r}
summary(additive.noLPA)
```

```{r}
additive.noLPA.noObesity <- lm(CANCER_CrudePrev ~ BINGE_CrudePrev + CSMOKING_CrudePrev + SLEEP_CrudePrev, data = df.calcs)

summary(additive.noLPA.noObesity)
plot(additive.noLPA.noObesity)
anova(additive.noLPA, additive.noLPA.noObesity)
```

##Interactions
```{r}
interactions.noLPA.noObesity <- lm(CANCER_CrudePrev ~ (BINGE_CrudePrev + CSMOKING_CrudePrev + SLEEP_CrudePrev)^2, data = df.calcs)

summary(interactions.noLPA.noObesity)
plot(interactions.noLPA.noObesity)
bptest(interactions.noLPA.noObesity)
```


```{r}
bc.NoLPA.NoObesity = boxcox(interactions.noLPA.noObesity,lambda=seq(-1,1))

bestlambda=bc.NoLPA.NoObesity$x[which(bc.NoLPA.NoObesity$y==max(bc$y))]
bestlambda
```

since the lambda is approzimately 0.5, we should try a sqrt transformation


## Square Root Transformation
```{r}
sqrt.noLPA.noObesity <- lm(sqrt(CANCER_CrudePrev) ~ (BINGE_CrudePrev + CSMOKING_CrudePrev + SLEEP_CrudePrev)^2, data = df.calcs)

summary(sqrt.noLPA.noObesity)
plot(sqrt.noLPA.noObesity)
bptest(sqrt.noLPA.noObesity)
```
















